A Dearth of Young Stars and Planets in the Kepler Field

### 2024-06-03 paper/ms.tex
further table formatting

\startlongtable
\begin{deluxetable*}{llllllllrrrrrrl}
  \tabletypesize{\scriptsize}
  \tablecaption{Ages of Kepler planets and planet candidates.  This
  version of the table is truncated to include only the youngest
  confirmed Kepler planets, sorted by the minimum of either the
  rotation or lithium-based age.  The full machine-readable table
  contains ages and age limits for \nnonfopkoissomeageinfo\ non-false
  positive KOIs with MES$>$10.  
  A bash script to decode any given $Q_{\rm star}$ quality flag is
  {\bf
  \href{https://gist.github.com/lgbouma/20368253f1a98da1b39cf32fdda0be13}{available
  online}}.  A python script to select stars with specific bit flags
  is {\bf \href{https://gist.github.com/lgbouma/87ad8bde42625e766a5a8857cc5a183a}{also available}}.
  \label{tab:planets}}
  %\toprule
  %\midrule
  %\endhead
  \startdata
  KOI & Kepler &  $T_{\rm eff}$ & $P_{\rm rot}$ & EW$_{\rm Li}^{\rm \ast}$ &
  $t_{\rm gyro}$ & $t_{\rm Li} $ & Consistent? & $R_{\rm p}$ & $P_{\rm orb}$ &
  $Q_{\rm planet}$ & $Q_{\rm star}$ &  Spec? & Comment \\
  -- &   -- & K & days & m\AA & Myr & Myr &  str &   Earths &    days &       int  & int & bool & -- \\
  \hline
  \input{table_subgyr_confirmed} %\\
% \rule
%  \input{table_subgyr_candidate}
  \enddata
  \tablecomments{
  	EW$_{\rm Li}^{\rm \ast}$ is the lithium equivalent width {\it after} subtracting a constant 7.5\,m\AA\ to account for the \ion{Fe}{1} 6707.44\,\AA\ blend (see Section~\ref{subsec:lithiumsel}).
  	Two checkmarks ($\checkmark\checkmark$) denote systems with two-sided $t_{\rm
  gyro}$ and $t_{\rm Li}$ for which both the age and the planet are expected to be reliable.  One checkmark ($\checkmark$) denotes two-sided $t_{\rm gyro}$ only. 
  ``Confirmed'' planets appear in the machine-readable
  version before``candidate'' planets.  The bit quality flags for the
  rotation-based ages, $Q_{\rm star}$, are described in
  Section~\ref{subsec:flags}.  Concisely summarized, they are:
  Bit 0: $T_{\rm eff}/{\rm K} \in [ 3800-6200]$ ?
  Bit 1: $\log g$$<$4.2?
  Bit 2: $M_{\rm G}$<3.9 or $M_{\rm G}>8.5$?
  Bit 3: In KEBC?
  Bit 4: Large $d_{\rm xm,Kep-Gaia}$?
  Bit 5: Confused Kep-Gaia crossmatch?
  Bit 6: Gaia DR3 non-single star?
  Bit 7: RUWE$>$1.4?
  Bit 8: Crowded?
  Bit 9: Far from main sequence?
  Bit 10: \citetalias{Santos_2021} CP/CB?
  Bit 11: $P_{\rm rot}$ not in the homogeneous
  \citetalias{Santos_2019} \citetalias{Santos_2021} sample?
  As an example, Kepler-1627 has $Q_{\rm star}$ flagged with bit 10 and bit 7.
  The analogous planet quality bitmask, $Q_{\rm planet}$, has the
  following meaning.
  Bit 0: Candidate not reliable?
  Bit 1: MES$<$10?
  Bit 2: Grazing?
  For a star to have a high likelihood of being ``reliable for
  gyrochronology'' we suggest a $Q_{\rm star}$ bitmask with bits 0--9
  not raised, and for a planet to be ``reliable'', we suggest
  $Q_{\rm planet}$ to be zero.
  }
\end{deluxetable*}

@@

\appendix
\section{Defining An ``Isochronally Too Far From The Main Sequence'' Boundary}
\label{app:linemethod}

@@

\begin{figure*}[!t]
  \begin{center}
    \leavevmode
        \includegraphics[width=0.48\textwidth]{gyroage_vs_teff_errs_linear.png}
    \includegraphics[width=0.48\textwidth]{gyroage_vs_teff_errs_showplanets_linear.png}
  \end{center}
  \vspace{-0.6cm}
  \caption{
    {\bf Rotation-based age as a function of effective temperature}.
    The entire Kepler sample is shown on the left.
    The KOI hosts are shown on the right.
    Each bar shows the $\pm$1$\sigma$ uncertainty for a star's age as
    derived from its rotation period.  
    Individual features are discussed in
    Appendix~\ref{app:age_diagnostic}.
    \label{fig:gyroage_vs_teff}
  }
\end{figure*}

>>>>

\startlongtable
\begin{deluxetable*}{llllllllrrrrrrl}
  \tabletypesize{\scriptsize}
  \tablecaption{Ages of Kepler planets and planet candidates.  This
  version of the table is truncated to include the youngest
  confirmed Kepler planets, sorted by the minimum of either the
  rotation or lithium-based age.  The full machine-readable table
  contains ages and age limits for \nnonfopkoissomeageinfo\ non-false
  positive KOIs with MES$>$10.  
  A bash script to decode the $Q_{\rm star}$ quality flag is
  {\bf
  \href{https://gist.github.com/lgbouma/20368253f1a98da1b39cf32fdda0be13}{available
  online}}.  A python script to select stars with specific bit flags
  is {\bf \href{https://gist.github.com/lgbouma/87ad8bde42625e766a5a8857cc5a183a}{also available}}.
  All quoted uncertainties are statistical; they do not incorporate
  possible systematic uncertainties, e.g.~in the absolute ages of open clusters.
  \label{tab:planets}}
  %\toprule
  %\midrule
  %\endhead
  \startdata
  KOI & Kepler &  $T_{\rm eff}$ & $P_{\rm rot}$ & EW$_{\rm Li}^{\rm \ast}$ &
  $t_{\rm gyro}$ & $t_{\rm Li} $ & Consistent? & $R_{\rm p}$ & $P_{\rm orb}$ &
  $Q_{\rm planet}$ & $Q_{\rm star}$ &  Spec? & Comment \\
  -- &   -- & K & days & m\AA & Myr & Myr &  str &   Earths &    days &       int  & int & bool & -- \\
  \hline
  \input{table_subgyr_confirmed} %\\
% \rule
%  \input{table_subgyr_candidate}
  \enddata
  \tablecomments{
  	EW$_{\rm Li}^{\rm \ast}$ is the lithium equivalent width {\it after} subtracting a constant 7.5\,m\AA\ to account for the \ion{Fe}{1} 6707.44\,\AA\ blend (see Section~\ref{subsec:lithiumsel}).
  	Two checkmarks ($\checkmark\checkmark$) denote systems with two-sided $t_{\rm
  gyro}$ and $t_{\rm Li}$ for which both the age and the planet are expected to be reliable.  One checkmark ($\checkmark$) denotes two-sided $t_{\rm gyro}$ only. 
  ``Confirmed'' planets appear in the machine-readable
  version before``candidate'' planets.  The bit quality flags for the
  rotation-based ages, $Q_{\rm star}$, are described in
  Section~\ref{subsec:flags}.  Concisely summarized, they are:
  Bit 0: $T_{\rm eff}/{\rm K} \in [ 3800-6200]$ ?
  Bit 1: $\log g$$<$4.2?
  Bit 2: $M_{\rm G}$<3.9 or $M_{\rm G}>8.5$?
  Bit 3: In KEBC?
  Bit 4: Large $d_{\rm xm,Kep-Gaia}$?
  Bit 5: Confused Kep-Gaia crossmatch?
  Bit 6: Gaia DR3 non-single star?
  Bit 7: RUWE$>$1.4?
  Bit 8: Crowded?
  Bit 9: Far from main sequence?
  Bit 10: \citetalias{Santos_2021} CP/CB?
  Bit 11: $P_{\rm rot}$ not in the homogeneous
  \citetalias{Santos_2019} \citetalias{Santos_2021} sample?
  As an example, Kepler-1627 has $Q_{\rm star}$ flagged with bit 10 and bit 7.
  The analogous planet quality bitmask, $Q_{\rm planet}$, has the
  following meaning.
  Bit 0: Candidate not reliable?
  Bit 1: MES$<$10?
  Bit 2: Grazing?
  For a star to have a high likelihood of being ``reliable for
  gyrochronology'' we suggest a $Q_{\rm star}$ bitmask with bits 0--9
  not raised, and for a planet to be ``reliable'', we suggest
  $Q_{\rm planet}$ to be zero.
  }
\end{deluxetable*}

@@

\appendix
\section{Defining An ``Isochronally Far From The Main Sequence'' Boundary}
\label{app:linemethod}

@@

\begin{figure*}[!t]
  \begin{center}
    \leavevmode
        \includegraphics[width=0.48\textwidth]{gyroage_vs_teff_errs_linear.png}
    \includegraphics[width=0.48\textwidth]{gyroage_vs_teff_errs_showplanets_linear.png}
  \end{center}
  \vspace{-0.6cm}
  \caption{
    {\bf Rotation-based age vs.~effective temperature}.
    The entire Kepler sample is shown on the left.
    KOI hosts are shown on the right.
    Each bar shows the $\pm$1$\sigma$ uncertainty for a star's age as
    derived from its rotation period.  
    Notable features are discussed in
    Appendix~\ref{app:age_diagnostic}.
    \label{fig:gyroage_vs_teff}
  }
\end{figure*}
